---
title: "Homework 2"
author: "Daniel Ojerant"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```


Read data
```{r}
trashwheel.df=
  read_xlsx("/Users/danie/Documents/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = cell_cols("A:N")) %>% janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)
)
```

Read precipitation

```{r}
 precip.2018=
  read_xlsx("/Users/danie/Documents/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation",
  skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
  
  precip.2017 =
  read_xlsx("/Users/danie/Documents/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2017 Precipitation",
  skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation
```{r}
month.df =
  tibble(
    month = 1:12,
    month_name = month.name
  )
precip.df =
  bind_rows(precip.2018,precip.2017)

left_join(precip.df, month.df, by = "month")
  
```

this dataset contains information from Mr.Trashwheel trash collector in Baltimore, Maryland. As treash enters the inner harbor, the trashwheel collects the trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel.df)` rows in our final dataset. Additional data sheets include month precipitation date


