---
title: "Homework 2"
author: "Daniel Ojerant"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```


Read data
```{r}
trashwheel.df=
  read_xlsx("/Users/danie/Documents/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = cell_cols("A:N")) %>% janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)
)
```

Read precipitation

```{r}
 precip.2018=
  read_xlsx("/Users/danie/Documents/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation",
  skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
  
  precip.2017 =
  read_xlsx("/Users/danie/Documents/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2017 Precipitation",
  skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation
```{r}
month.df =
  tibble(
    month = 1:12,
    month_name = month.name
  )
precip.df =
  bind_rows(precip.2018,precip.2017)

left_join(precip.df, month.df, by = "month")
  
```

This dataset contains information from Mr.Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects the trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel.df)` rows in our final dataset. Additional data sheets include month precipitation date

## Problem 2

```{r, include=FALSE}
NYC.df = 
  read_csv("/Users/danie/Documents/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entrance_type, entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = 1, "NO" = 0))
```

This data set is 1868 x 19  with variables including the line, station name, station latitude, station longitude, the 11 routes, the entrance type, the entry, vending, and ADA approval. I cleaned the dataset by only importing those following variables while also recoding the entry variable from character to logical.This data set is 1868 x 19. This data isn't tidy because all the route variables can be represented as one single variable.  

```{r, include=FALSE}

Distinct.NYC.df = distinct(NYC.df, line, station_name)
```
There is a total number of `r nrow(distinct(NYC.df, line, station_name))` distinct stations.

```{r, include=FALSE}
NYC1.df = 
  filter(NYC.df, ada == "TRUE") %>% 
  distinct(line, station_name)
```
 `r nrow(distinct(NYC1.df, line, station_name))` of the `r nrow(distinct(NYC.df, line, station_name))` distinct stations are ADA complaint.


```{r, include=FALSE}
transit.df = 
  read_csv("/Users/danie/Documents/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entrance_type, entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = 1, "NO" = 0)) %>% 
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  ) 

```


```{r,include=FALSE}
A.Line = filter(transit.df, route_name == "A",) %>% 
  distinct(line, station_name)

A.Line.ada = filter(transit.df, route_name == "A", ada == "TRUE") %>% 
  distinct(line, station_name)
```

There are `r nrow(A.Line)` distinct stations that serve on the A line which `r nrow(A.Line.ada)` are ADA complient