---
title: "p8105_hw2_do2381"
author: "Daniel Ojerant"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```


## Problem 1
```{r, include=FALSE}
trashwheel.df=
  read_xlsx("~/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = cell_cols("A:N")) %>% janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)
) 
  
```


```{r, include=FALSE}
 precip.2018=
  read_xlsx("~/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation",
  skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
  
  precip.2017 =
  read_xlsx("~/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2017 Precipitation",
  skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```



```{r, include=FALSE}
month.df =
  tibble(
    month = 1:12,
    month_name = month.name
  )
precip.df =
  bind_rows(precip.2018,precip.2017)

left_join(precip.df, month.df, by = "month")
  
```

This dataset contains information from Mr.Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects the trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel.df)` rows in our final dataset. 

## Problem 2

```{r, include=FALSE}
NYC.df = 
  read_csv("~/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entrance_type, entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = 1, "NO" = 0))
```

This data set is 1868 x 19  with variables including the line, station name, station latitude, station longitude, the 11 routes, the entrance type, the entry, vending, and ADA approval. I cleaned the dataset by only importing those following variables while also recoding the entry variable from character to logical.This data set is 1868 x 19. This data isn't tidy because all the route variables can be represented as one single variable.  

```{r, include=FALSE}

Distinct.NYC.df = distinct(NYC.df, line, station_name)
```
There is a total number of `r nrow(distinct(NYC.df, line, station_name))` distinct stations.

```{r, include=FALSE}
NYC1.df = 
  filter(NYC.df, ada == "TRUE") %>% 
  distinct(line, station_name)
```
Of the `r nrow(distinct(NYC.df, line, station_name))` distinct stations, `r nrow(distinct(NYC1.df, line, station_name))` are ADA complaint.


```{r, include=FALSE}
transit.df = 
  read_csv("~/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entrance_type, entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = 1, "NO" = 0)) %>% 
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  ) 

```


```{r,include=FALSE}
A.Line = filter(transit.df, route_name == "A") %>% 
  distinct(line, station_name)

A.Line.ada = filter(transit.df, route_name == "A", ada == "TRUE") %>% 
  distinct(line, station_name)
```

There are `r nrow(A.Line)` distinct stations that serve on the A line which `r nrow(A.Line.ada)` are ADA complaint.

```{r, include = FALSE }
entry.yes.df =
  transit.df %>% 
  filter(vending == "NO", entry == 1 ) %>% 
  count()

vending.no.df =
   transit.df %>% 
  filter(vending == "NO") %>% 
  count()
```

The proportion of station of entrances / exits that does not have vending is `r entry.yes.df ` / `r vending.no.df `

## Problem 3

```{r, include=FALSE}
pols.df = 
  read_csv("~/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, c("year","month","day"), sep = "-") %>%
  mutate(president = prez_gop + prez_dem, month = recode( month, "01" = "1","02" = "2","03" = "3","04" = "4","05" = "5","06" = "6","07" = "7","08" = "8","09" = "9")) %>% 
  select( -prez_gop, -prez_dem, -day)
  
snp.df = 
  read_csv("~/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>%
  separate(date, c("month","day","year"), sep = "/") %>% 
  select(-day) %>% 
  relocate(year, month)

unemployment.df = 
  read_csv("~/Columbia Semester 1 Files/Data Science  R Code/Data Wrangling/p8105_hw2_do2381/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>%
  mutate(year = as.character(year)) %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percent"
  ) %>% 
 mutate(month = recode(month, "jan"="January","feb"="Febuary","mar"="March","apr"="April","may"="May","jun"="June","jul"="July","aug"="August","sep"="September","oct"="October","nov" ="November","dec"="December"))
 
```
The pols dataset has `r nrow(pols.df)` rows and `r ncol(pols.df)` columns. This dataset includes variables like the month, year and number of national politicians who are democratic or republican at every given year. The years range from 1947-2015.

The s&p data has `r nrow(snp.df)` rows and `r ncol(snp.df)` columns. This dataset includes the variables year, month, and close.
The years in this dataset range from 1950-2015.

The unemployment dataset has `r nrow(unemployment.df)` rows and `r ncol(unemployment.df)` columns. This dataset includes the variables year, month, and percent. The years in this dataset range from 1948-2015.

```{r, include=FALSE}
bind.df =
  left_join(pols.df, snp.df) %>% 
  transform(month = as.numeric(month))

 pols.snp.df = 
   left_join(bind.df, month.df) %>% 
   relocate(month_name) %>% 
   select(-month, month = month_name) 
   
  pols.snp.unemployment.df = 
   left_join(pols.snp.df, unemployment.df)
```

The final merged datset has `r nrow(pols.snp.unemployment.df)` rows and `r ncol(pols.snp.unemployment.df)` columns. This dataset contains the data from all 3 datasets. ranging from 1947-2015.


